Personal Total Income of Individual in the United States of America in 2023

### 2024-10-26 paper/paper.qmd
Review the result with updated data

| Metric          | Value            |
|-----------------|------------------|
| Best Model MSE  | 6,057,429,750.02 |
| Best Model RMSE | 77,829.49        |
:::

@@

| Feature         | Importance |
|-----------------|------------|
| cat\_\_EDUC_new | 0.337065   |
| cat\_\_IND1990  | 0.243454   |
| num\_\_AGE      | 0.193926   |
| cat\_\_SEX      | 0.149342   |
| cat\_\_MARST    | 0.055948   |
| cat\_\_MORTGAGE | 0.013064   |
| cat\_\_SCHLTYPE | 0.004251   |
| cat\_\_VETSTAT  | 0.002948   |
| cat\_\_OWNERSHP | 0.000000   |
:::

@@

| Metric                         | Value            |
|--------------------------------|------------------|
| Mean Squared Error (MSE)       | 5,191,183,577.52 |
| Root Mean Squared Error (RMSE) | 72,049.87        |
:::

@@

::: {#tbl-XGBoostCV}
\begin{center}
Table 7: XGBoost Cross-Validation Performance
\end{center}

| Metric                | Value     |
|-----------------------|-----------|
| Cross-Validation RMSE | 71,994.55 |
:::

@@

We employed a Random Forest model to predict income, tuning several hyperparameters to improve model performance. The best model was achieved with the following parameters: **max_depth=10**, **min_samples_leaf=4**, **min_samples_split=10**, and **n_estimators=300**. As we shown in Table \ref{tbl-randomforestMSE}, the final model had a Mean Squared Error (MSE) of 6,057,429,750 and a Root Mean Squared Error (RMSE) of 77,829.49, indicating a substantial improvement in prediction accuracy compared to the linear regression model.

@@

Other variables, such as `cat__SEX` (Sex) and `cat__MARST` (Marital Status), show moderate importance, while factors like `cat__MORTGAGE` and `cat__SCHLTYPE` (School Type) have much smaller impacts. Interestingly, `cat__OWNERSHP` (Home Ownership) has no contribution to the model, as its importance score is 0.000.

@@

The XGBoost model, a gradient boosting algorithm, was optimized using a hyperparameter grid search, yielding the best parameters: **learning_rate=0.1**, **max_depth=6**, **min_child_weight=5**, and **n_estimators=300**. These parameters were selected to balance model complexity and overfitting control. As we shown in Table \ref{tbl-XGBoostMSE}, the final model achieved a Mean Squared Error (MSE) of 5,191,183,577.52 and a Root Mean Squared Error (RMSE) of 72,049.87, making it the best-performing model among the three evaluated.

@@

The XGBoost model’s Cross-Validation RMSE (Table \ref{tbl-XGBoostCV}) was 71,994.55, which aligns closely with the final RMSE, indicating that the model generalizes well across different data subsets. This consistency in performance suggests that XGBoost effectively captures the relationship between predictors and income without overfitting.

>>>>

| Metric          | Value            |
|-----------------|------------------|
| Best Model MSE  | 6,057,426,411.65 |
| Best Model RMSE | 77,829.47        |
:::

@@

| Feature         | Importance |
|-----------------|------------|
| cat\_\_EDUC_new | 0.337065   |
| cat\_\_IND1990  | 0.243454   |
| num\_\_AGE      | 0.193926   |
| cat\_\_SEX      | 0.149342   |
| cat\_\_MARST    | 0.055948   |
| cat\_\_MORTGAGE | 0.013064   |
| cat\_\_SCHLTYPE | 0.004251   |
| cat\_\_VETSTAT  | 0.002948   |
:::

@@

| Metric                         | Value            |
|--------------------------------|------------------|
| Mean Squared Error (MSE)       | 5,854,629,848.13 |
| Root Mean Squared Error (RMSE) | 76,515.55        |
:::

@@

::: {#tbl-XGBoostCV}
\begin{center}
Table 7: Cross-Validation RMSE for XGBoost Model
\end{center}

| Fold  | RMSE       |
|-------|------------|
| 1     | 76,693.86  |
| 2     | 75,692.41  |
| 3     | 76,193.42  |
| 4     | 77,365.82  |
| 5     | 76,552.10  |
| **Average** | **76,499.52** |
:::

@@

We employed a Random Forest model to predict income, tuning several hyperparameters to improve model performance. The best model was achieved with the following parameters: **max_depth=10**, **min_samples_leaf=4**, **min_samples_split=10**, and **n_estimators=300**. As we shown in Table \ref{tbl-randomforestMSE}, the final model had a Mean Squared Error (MSE) of 6,057,426,411.65 and a Root Mean Squared Error (RMSE) of 77,829.47, indicating a substantial improvement in prediction accuracy compared to the linear regression model.

@@

Other variables, such as `cat__SEX` (Sex) and `cat__MARST` (Marital Status), show moderate importance, while factors like `cat__MORTGAGE` and `cat__SCHLTYPE` (School Type) have much smaller impacts.

@@

The XGBoost model, a gradient boosting algorithm, was optimized using a hyperparameter grid search, yielding the best parameters: **learning_rate=0.1**, **max_depth=6**, **min_child_weight=5**, and **n_estimators=300**. These parameters were selected to balance model complexity and overfitting control. As we shown in Table \ref{tbl-XGBoostMSE}, the final model achieved a Mean Squared Error (MSE) of 5,854,629,848.13 and a Root Mean Squared Error (RMSE) of 76,515.55, making it the best-performing model among the three evaluated.

@@

The XGBoost model’s Cross-Validation RMSE (Table \ref{tbl-XGBoostCV}) was 76,499.52, which aligns closely with the final RMSE, indicating that the model generalizes well across different data subsets. This consistency in performance suggests that XGBoost effectively captures the relationship between predictors and income without overfitting.
